---
title: "PCA - Homework 4 - Lecture 6"
author: "Aleksey Kramer"
date: "November 14, 2015"
output: html_document
---
#### After conduction Principle Component Analysis (as shown in the step-by-step manner below), components 1, 2, and 3 appear to explain 99% of the variance in the data set.

#### Set working directory and load the dataset in the memory
```{r}
# Set working directory
setwd('C:\\Users\\Aleksey\\Documents\\School\\UW_Data_Science\\UW_Data_Science_450\\Week6\\homework')

# Read CSV file
data = read.csv("women_TnF_1984.csv")
```

#### Look at the structure of the data
```{r}
# investigate data frame
str(data)
```

#### Visually examine the dataset and remove unique identifier
```{r}
# look at the data
head(data)

# Discovered unique identifier data$Obs
# Remove unique identifier
data$Obs = NULL
```

#### Visually examine the dataset again and convert minutes to seconds for fields m800, m1500, m3000, and Marathon
```{r}
# display common stats
summary(data)

# display first 10 rows
head(data)

# Long way to convert data to seconds to be normalized
data$m800 = data$m800 * 60 
data$m1500 = data$m1500 * 60
data$m3000 = data$m3000 * 60 
data$Marathon = data$Marathon * 60
```

#### Visually inspect the data again.
```{r}
# display first 10 rows
head(data)

# display common stats again on the transformed data
summary(data)
```

#### Naming rows in data to be country names from data[,1] and removing 'Country' field from the dataset
```{r}

rownames(data) = data[,1]
data[,1] = NULL
```

#### Visually inspect the datset again
```{r}
# display first 10 rows to make sure Country column is removed
head(data)
```

#### Perform Principal Component Analysis (PCA) and display its summary
```{r}
# Running PCA
fit <- princomp(data, cor=TRUE)

# Displaying summary of the fit
summary(fit)
```

#### Display Loadings
```{r}
# Display loadings
loadings(fit)
```

#### Plot the 'fit' obtained by running PCA. Plot clealy shows componnents 1, 2, and 3 to be the principal components derived (explain more than 99% of the variance in the dataset as seen in the display of loadings)
```{r}
# Plot the fit
plot(fit, type="l")
```

#### Display the scores for each record
```{r}
# Display scores
fit$scores
```

#### Create a biplot to visualize the directions of the most varying data (in 2D)
```{r}
# Display biplot
biplot(fit)
```

#### End of PCA assignment